--- 
title: "Chapter_1" 
author: "Philippe Glaziou, Pete Dodd" 
date: "`r Sys.Date()`" 
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
--- 


```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(
  echo = FALSE,
  results = "asis",
  message = FALSE,
  warning = FALSE
)
library(data.table)
library(gtbreport)
library(ggpubr)
library(here)

load(here('report/data/gtb.rda'))


``` 






`r anch("Fig. 1.1")`

### <span style="color:#F21905">Fig. 1.1</span> Global trend in case notifications of people newly diagnosed with TB, 2016&#8211;2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.1, fig.alt="Global trend in notifications of new and relapse TB cases, 2016-2020"} 

qplot(
  year,
  c.newinc / 1e6,
  data = global[year >= 2016],
  geom = 'line',
  size = I(1)
) +
  xlab('') + ylab('Notifications per year (millions)') +
  theme_gtb()

```









`r anch("Fig. 1.2")`

### <span style="color:#F21905">Fig. 1.2</span> Trends in notifications of new and relapse TB cases by WHO region, 2016&#8211;2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2, fig.alt="Trends in notifications of new and relapse TB cases by WHO region, 2016-2020"} 

qplot(
  year,
  c.newinc / 1e6,
  data = regional[year >= 2016],
  geom = 'line',
  size = I(1),
  colour = I('dark blue')
) +
  xlab('') + ylab('Notifications per year (millions)') +
  facet_wrap( ~ region, scales = 'free_y') +
  theme_gtb()

```









`r anch("Fig. 1.3")`

### <span style="color:#F21905">Fig. 1.3</span> Trends in monthly or quarterly notifications of TB cases in selected high TB burden countries, January 2020&#8211;June 2021, ^a^
<span class="subhead">The black line indicates the average number of monthly or quarterly case notifications in 2019.</span>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8, fig_1.3, fig.alt="Trends in monthly or quarterly notifications of TB cases from January 2020 to June 2021, selected high TB burden countries"} 

qplot(time,
      value,
      data = monthly[(year == 2020 &
                        nm == 12) |
                       (year == 2021 & variable <7 & nm > 0)],
      geom = 'line',
      size = I(1),
colour = I('#008000') )+
  geom_hline(
    aes(yintercept = n2019 / 12),
    linetype = I(1),
    colour = I('#000000'),
    size = I(1)
  ) +
  scale_x_date(date_breaks = "2 months", date_labels = "%m") +
  expand_limits(x = as.Date('20-12-2019', format = '%d-%m-%Y')) +
  facet_wrap(~ country, scales = 'free_y') +
  xlab('Month, 2020-2021') + ylab('Notifications per month') +
  theme_gtb()

qplot(variable,
      value,
      data = qty[year == 2020],
      geom = 'line',
      size = I(1),
     colour = I('#008000'),) +
  geom_hline(
    aes(yintercept = n2019 / 4),
    linetype = I(1),
    colour = I('#000000'),
    size = I(1)
  ) +
  facet_wrap( ~ country, scales = 'free_y') +
  xlab('Quarter, 2020') + ylab('Notifications per quarter') +
  theme_gtb()

```

<div class="footnote"> ^a^ Data are shown for countries that were able to report provisional national numbers for all months or quarters to WHO by July 2021.</div> 








`r anch("Fig. 1.4")`

### <span style="color:#F21905">Fig. 1.4</span> Trends in monthly notifications of TB cases in India, January 2020–June 2021^a^

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8, fig_1.4, fig.alt="Trends in monthly TB case notifications in India in 2020 and 2021"} 

qplot(
  month,
  (public + private) / 1000,
  data = ind,
  geom = 'line',
  size = I(1)
) +
  geom_hline(aes(yintercept = n2019 / 1000),
             colour = I('#3770c9'),
             size = I(1)) +
  geom_line(aes(month, private / 1000),
            colour = I('red'),
            size = I(1)) +
  ylab('Notifications per month (thousands)') +
  xlab('Month, 2020-2021') +
  scale_x_date(date_breaks = "1 month", date_labels = "%m/%y") +
  scale_y_continuous(breaks = seq(50, 200, by = 25)) +
  annotate(
    'text',
    x = as.Date('2020-07-01'),
    y = 190,
    label = 'Average case notifications in 2019',
    colour = I('#3770c9')
  ) +
  annotate('text',
           x = as.Date('2020-07-01'),
           y = 120,
           label = 'Total case notifications') +
  annotate(
    'text',
    x = as.Date('2020-07-01'),
    y = 60,
    label = 'Notifications from the private sector',
    colour = I('red')
  ) +
  theme_gtb()

```

<div class="footnote">^a^ These data were accessed from https://reports.nikshay.in/Reports/TBNotification on 20 July 2021.</div> 








`r anch("Fig. 1.5")`

### <span style="color:#F21905">Fig. 1.5</span> TB notifications in 2020 compared with 2019 in the 30 high TB burden countries

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.5, fig.alt="Shortfaill in TB notifications in the 30 high TB burden countries in 2020 compared with 2019"} 

dta <- est[year==2020 & g.hbc==TRUE, .(iso3, country, shortfall = short * 100)]

ggbarplot(
  dta,
  x = 'country',
  y = 'shortfall',
  fill = "#00aaad",
  color = 'white',
  sort.val = "desc",
  xlab = '',
  ylab = 'Percentage of TB notifications in 2020 compared with 2019'
) + 
  rotate() + 
  geom_hline(yintercept=100) + 
  scale_y_continuous(breaks=seq(0,110, by=10))

```










`r anch("Fig. 1.6")`

### <span style="color:#F21905">Fig. 1.6</span> The 16 countries with the largest contributions to the global shortfall in TB notifications in 2020 compared with 2019

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.6, fig.alt="The 16 countries with the largest contributions to the global shortfall in TB notifications in 2020 compared with 2019"} 

gshort <- global[year==2019, c.newinc] - global[year==2020, c.newinc]
dta <-
  est[year >= yr-1 &
        iso3 %in% unique(model$iso3), .(shortfall = -diff(c.newinc)), by=.(iso3, country)]
dta[, pshort := 100 * shortfall / gshort]

# ggplot(data = dta, aes(reorder(country, shortfall), shortfall)) +
#   geom_bar(stat = 'identity',
#            fill = I('#00aaad'),
#            colour = I('black')) +
#   xlab('')  + ylab('Millions (2019)') +
#   coord_flip() +
#   theme_gtb()

ggbarplot(
  dta,
  x = 'country',
  y = 'pshort',
  fill = "#00aaad",
  color = 'white',
  sort.val = "asc",
  xlab = '',
  ylab = 'Percentage of global shortfall in 2020 TB notifications compared with 2019'
) + 
  rotate()

```









`r anch("Fig. 1.7")`

### <span style="color:#F21905">Fig. 1.7</span> Estimated impact of the COVID-19 pandemic on TB mortality for 16 selected countries, up to 2025

<span class="subhead">Standardized TB mortality rate (including HIV)^a^</span> 


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=11, fig_1.7, fig.alt="Estimated impact of the COVID-19 pandemic on TB mortality for 16 selected countries, up to 2025"} 


qplot(
  as.Date(date),
  rebest * 100,
  data = model[measure == 'mort' &
                 hiv == 'a'],
  colour = scenario,
  size = I(1),
  geom = 'line'
) +
  xlab('') + ylab('TB mortality change relative to January 2020 (%)') +
  # expand_limits(y = 0) +
  facet_wrap( ~ country, scales = 'free_y') +
  #scale_color_brewer(palette = 'Set1') +
  scale_colour_manual(values = c("black", "red"))+
#  labs(title = bquote("Standardized TB mortality rate including" ~HIV^a)) +
  theme_gtb() +
  theme(legend.position = 'none')

```

<div class="footnote"> ^a^ These estimates are standardized so that rates in January 2020 equal 100 and all subsequent rates are relative to January 2020. For example, a reading of 115 translates into a 15% increase relative to January 2020. Baseline is a scenario of no COVID-19 disruptions based on pre-2020 trends. The impact of COVID-19 related disruptions on estimated mortality is noticeable from 2020 onward.</div> 







`r anch("Fig. 1.8")`

### <span style="color:#F21905">Fig. 1.8</span> Estimated impact of the COVID-19 pandemic on TB incidence for 16 selected countries, up to 2025

<span class="subhead">Standardized TB incidence rate^a^</span> 


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=11, fig_1.8, fig.alt="Estimated impact of the COVID-19 pandemic on TB incidence for 16 selected countries, up to 2025"} 

qplot(
  as.Date(date),
  rebest * 100,
  data = model[measure == 'inc' &
                 hiv == 'a'],
  colour = scenario,
  size = I(1),
  geom = 'line'
) +
  xlab('') + ylab('TB incidence change relative to January 2020 (%)') +
  facet_wrap( ~ country) +
  scale_colour_manual(values = c("black", "blue"))+
#  labs(title = bquote('Standardized TB incidence' ~rate^a)) +
  theme_gtb() +
  theme(legend.position = 'none')

```

<div class="footnote">^a^ These estimates are standardized so that rates in January 2020 equal 100 and all subsequent rates are relative to January 2020. For example, a reading of 115 translates into a 15% increase relative to January 2020. Baseline is a counterfactual scenario of no COVID-19, based on pre-2020 trends. The impact of COVID-19 related disruptions on estimated incidence is limited in 2020 and more noticeable in subsequent years.</div> 







