--- 
title: "Chapter_2" 
author: "Philippe Glaziou, Pete Dodd" 
date: "`r Sys.Date()`" 
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
--- 



```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(echo = FALSE,  
                      results = "asis", 
                      message = FALSE, 
                      warning = FALSE) 
library(data.table)
library(gtbreport)
library(here)
library(kableExtra)
library(gridExtra)
library(whomap)
library(ggpubr)
library(productplots)
library(RColorBrewer)

load(here('report/data/gtb.rda'))

``` 







`r anch("Table 2.2.1")`

### <span style="color:#F21905">Table 2.2.1</span> Global and regional estimates of TB mortality, numbers (in thousands) and rates (per 100 000 population) in 2020
```{r tab_2.2.1}

tab.header2 <- c('Region or country group','Best estimate','Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High'
                  )

knitr::kable(cbind(tab1[c(32:37, 31, 38), c(1, 9:14)], tab1b[c(32:37, 31, 38), 8:13]),
             format = "html",
             align = 'rcccccccccccc',
             col.names = tab.header2,
             # Add a table ID so that it can be styled using extra CSS in Sitefinity
             table.attr = "id='burden_num_table'") |>
add_header_above(header = c(" " = 1, "HIV-negative" = 3, "HIV-positive" = 3, "HIV-negative" = 3, "HIV-positive" = 3)) |>
add_header_above(header = c(" " = 1, "Number of deaths (in thousands)" = 6, "Rate per 100 000 population" = 6)) 
  # pack_rows("High burden countries", 1, 30) |>
  # pack_rows('Groupings', 31, 38) 

```
<div class="footnote">
Low and high are the 5th and 95th percentiles of the uncertainty interval (UI)
</div>






`r anch("Fig. 2.2.1")`

### <span style="color:#F21905">Fig. 2.2.1</span> Global trends in the estimated number of TB deaths (left) and the mortality rate (right), 2000&#8211;2020 
<span class="subhead">Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.2.1, fig.alt="Global trends in the estimated number of TB deaths (left) and the mortality rate (right), 2000-2020"} 

p1 <-
  qplot(
    year,
    mort.num / 1e6,
    data = global,
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e6, ymax = mort.hi.num / 1e6),
    fill = I('grey40'),
    alpha = 0.4
  ) +
  ylab('Millions per year (log scale)') + xlab('') +
  geom_line(aes(year, mort.h.num / 1e6), colour = I('lightskyblue2')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo.num / 1e6, ymax = mort.h.hi.num / 1e6),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.nh.num / 1e6), colour = I('blue')) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo.num / 1e6, ymax = mort.nh.hi.num / 1e6),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.num / 1e6), colour = I('grey20')) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e6, ymax = mort.hi.num / 1e6),
    fill = I('grey40'),
    alpha = 0.4
  ) +
  annotate(
    'text',
    y = 1.15,
    x = 2000,
    label = '2020 milestone (total)',
    hjust = 0,
    size = I(4)
  ) +
  expand_limits(y = 0) +
  geom_hline(aes(yintercept = mort.num[16] * 0.65 / 1e6), linetype = I(2)) +
  annotate(
    'text',
    x = 2013,
    y = 2,
    label = 'Total',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 1.2,
    label = 'HIV-negative',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 0.20,
    label = 'HIV-positive',
    hjust = 0,
    size = I(4)
  ) +
  scale_y_log10(breaks=c(0.3,0.5,1,1.5,2)) +
  theme_gtb() + theme(legend.position = 'none')

p2 <-
  qplot(
    year,
    mort,
    data = global,
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(aes(year, ymin = mort.lo, ymax = mort.hi),
              fill = I('grey40'),
              alpha = 0.4) +
  ylab('Rate per 100 000 population per year (log scale)') + xlab('') +
  geom_line(aes(year, mort.h), colour = I('lightskyblue2')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo, ymax = mort.h.hi),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.nh), colour = I('blue')) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo, ymax = mort.nh.hi),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort), colour = I('grey20')) +
  geom_ribbon(aes(year, ymin = mort.lo, ymax = mort.hi),
              fill = I('grey40'),
              alpha = 0.4) +
  expand_limits(y = 0) +
  annotate(
    'text',
    x = 2013,
    y = 28,
    label = 'Total',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 15,
    label = 'HIV-negative',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 2,
    label = 'HIV-positive',
    hjust = 0,
    size = I(4)
  ) +
  scale_y_log10() +
  theme_gtb() + theme(legend.position = 'none')


mp <- grid.arrange(p1, p2, ncol = 2)

```
 





`r anch("Fig. 2.2.2")`

### <span style="color:#F21905">Fig. 2.2.2</span> Top causes of death worldwide in 2019^a,b^
<span class="subhead">Deaths from TB among HIV-positive people are shown in grey.</span>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=6, fig_2.2.2, fig.alt="Top causes of death worldwide in 2019"} 

tbhiv <- global$mort.h.num[global$year == 2019] / 1e3
top10$tbhiv <- c(rep(0, 12), tbhiv, rep(0, 7))
top10$n <- top10$deaths / 1000 + top10$tbhiv / 1000

ggplot(data = top10, aes(reorder(cause, deaths), n)) +
  geom_bar(stat = 'identity',
           fill = I('grey70'),
           colour = I('black')) +
  geom_bar(
    aes(cause, deaths / 1000),
    stat = 'identity',
    fill = I('#00aaad'),
    colour = I('black')
  ) +
  xlab('')  + ylab('Number of deaths (millions)') +
  scale_y_continuous(breaks = 0:9) +
  coord_flip() +
  theme_gtb()




```
<div class="footnote"> ^a^ This is the latest year for which estimates for all causes are currently available. See WHO estimates, available at https://www.who.int/data/gho/data/themes/mortality-and-global-health-estimates/ghe-leading-causes-of-death.<br>
^b^ Deaths from TB among HIV-positive people are officially classified as deaths caused by HIV/AIDS in the International Classification of Diseases. </div> 






`r anch("Fig. 2.2.3")`

### <span style="color:#F21905">Fig. 2.2.3</span> Estimated number of deaths from HIV/AIDS and TB in 2020^a,b^ 
<span class="subhead">Deaths from TB among HIV-positive people are shown in grey.</span>
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png',fig.height=6, fig_2.2.3, fig.alt="Estimated number of deaths from HIV/AIDS and TB in 2020"} 

hivd <- unaids[, .(sum(mort.hiv.num, na.rm = T)), by = year]
cod <- data.table(
  cause = c('HIV/AIDS', 'TB'),
  n = c((last(hivd$V1) - last(global$mort.h.num)) / 1e6, last(global$mort.nh.num) /
          1e6),
  tbhiv = rep(last(global$mort.h.num) / 1e6, 2)
)
cod[, total := n + tbhiv]

ggplot(data = cod, aes(reorder(cause, total), total)) +
  geom_bar(stat = 'identity',
           fill = I('grey70'),
           colour = I('black')) +
  geom_bar(
    aes(cause, n),
    stat = 'identity',
    fill = I('#00aaad'),
    colour = I('black')
  ) +
  xlab('') + ylab('Millions (2020)') +
  coord_flip() +
  theme_gtb()



```
<div class="footnote"> ^a^ For HIV/AIDS, the latest estimates of the number of deaths in 2020 that have been published by UNAIDS are available at http://www.unaids.org/en/ (accessed 16 August 2021). For TB, the estimates for 2020 are those published in this report.<br>
^b^ Deaths from TB among HIV-positive people are officially classified as deaths caused by HIV/AIDS in the International Classification of Diseases.</div> 





`r anch("Fig. 2.2.4")` 

### <span style="color:#F21905">Fig. 2.2.4</span> Global trends in the estimated number of deaths caused by TB and HIV (in millions), 2000&#8211;2020^a,b^
<span class="subhead">Shaded areas represent uncertainty intervals.</span> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.2.4, fig.alt="Global trends in the estimated number of deaths caused by TB and HIV (in millions), 2000–2020"} 

m <- 1e5
ghiv.mort <-
  est[, addXY(mort.hiv / m, r.sd = mort.hiv.sd / m, weights = as.numeric(pop)), by =
        year]
dta <- merge(global, ghiv.mort, by = 'year')
M <- 1e6

ggplot(data = dta, aes(year, mort.h.num / M)) +
  geom_line(colour = I('lightskyblue2')) +
  geom_line(aes(year, mort.nh.num / M), colour = I('blue')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo.num / M, ymax = mort.h.hi.num / M),
    fill = I("lightskyblue2"),
    alpha = I(.4)
  ) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo.num / M, ymax = mort.nh.hi.num / M),
    fill = I("blue"),
    alpha = I(.4)
  ) +
  geom_line(aes(year, r.num / M), colour = I('grey50')) +
  geom_ribbon(
    aes(year, ymin = r.lo.num / M, ymax = r.hi.num / M),
    fill = I("grey50"),
    alpha = I(.4)
  ) +
  xlab('') + ylab('Millions of deaths per year (log scale)') +
  annotate(
    'text',
    x = 2009,
    y = 1.9,
    label = 'HIV deaths',
    size = I(5),
    colour = I('grey50')
  ) +
  annotate(
    'text',
    x = 2005,
    y = 1.3,
    label = 'TB deaths in\nHIV-negative people',
    size = I(5),
    colour = I('blue')
  ) +
  annotate(
    'text',
    x = 2005,
    y = 0.30,
    label = 'TB deaths in\nHIV-positive people',
    size = I(5),
    colour = I('lightskyblue4')
  ) +
  expand_limits(y = 0) + 
  scale_y_log10(breaks=c(0.3,0.5,1,1.5,1.8)) +
  theme_gtb()


```
<div class="footnote"> ^a^ For HIV/AIDS, the latest estimates of the number of deaths in 2020 that have been published by UNAIDS are available at http://www.unaids.org/en/ (accessed 16 August 2021). For TB, the estimates for 2020 are those published in this report.<br>
^b^ Deaths from TB among HIV-positive people are officially classified as deaths caused by HIV/AIDS in the International Classification of Diseases.</div> 



`r anch("Fig. 2.2.5")`

### <span style="color:#F21905">Fig. 2.2.5</span> Trends in estimated TB mortality rates by WHO region, 2000&#8211;2020
<span class="subhead">Estimated TB mortality rates among HIV-negative people are shown in <span style="color:#3232ff">blue</span> and estimated mortality rates among HIV-positive people are shown in <span style="color:#87ceeb">light blue</span>. Shaded areas represent uncertainty intervals.</span> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height = 8, fig_2.2.5, fig.alt="Trends in estimated TB mortality rates by WHO region, 2000−2020"} 

qplot(
    year,
    mort.nh,
    data = regional,
    geom = 'line',
    colour = I('blue')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo, ymax = mort.nh.hi),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.h), colour = I('lightskyblue2')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo, ymax = mort.h.hi),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  expand_limits(y=0) +
  facet_wrap(~ region, scale = 'free_y') +
  xlab('') + ylab('Mortality rate per 100 000 population per year (log scale)') +
  #  scale_y_continuous(trans='log10', breaks=c(0.1,1,5,10,20,30,50)) +
  theme_gtb() +
  scale_y_log10() +
  theme(legend.position = 'none')


```





`r anch("Fig. 2.2.6")` 

### <span style="color:#F21905">Fig. 2.2.6</span> Trends in the estimated absolute number of TB deaths (HIV-positive and HIV-negative, in thousands) by WHO region, 2000&#8211;2020 
<span class="subhead">Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height = 8, fig_2.2.6, fig.alt="Trends in the estimated absolute number of TB deaths (HIV-positive and HIV-negative, in thousands) by WHO region, 2000-2020"} 

qplot(
    year,
    mort.num / 1e3,
    data = regional,
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_hline(aes(yintercept = mort.milestone / 1e3), linetype = I(2)) +
  expand_limits(y=0) +
  facet_wrap(~ region, scale = 'free_y') +
  xlab('') + ylab('Total TB deaths (thousands, log scale)') +
  scale_y_log10() +
  theme_gtb() +
  theme(legend.position = 'none')


```







`r anch("Fig. 2.2.7a")` 

### <span style="color:#F21905">Fig. 2.2.7a</span> Trends in the estimated absolute number (in thousands) of TB deaths (HIV-positive and HIV-negative TB) in the 30 high TB burden countries, 2000&#8211;2020
<span class="subhead">Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=12, fig_2.2.7a, fig.alt="Trends in the estimated absolute number of TB deaths (HIV-positive and HIV-negative TB) in the 30 high TB burden countries, 2000−2020"} 
  qplot(
    year,
    mort.num / 1e3,
    data = hest,
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_hline(aes(yintercept = mort.milestone / 1e3), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
    facet_wrap(~country, nrow=6, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
  xlab('') + ylab('TB deaths (total, in thousands) per year (log scale)') +
  expand_limits(y = 0) +
  scale_y_log10() +
  theme_gtb()


```






`r anch("Fig. 2.2.7b")`

### <span style="color:#F21905">Fig. 2.2.7b</span> Trends in the estimated absolute number (in thousands) of TB deaths (HIV-positive and HIV-negative TB) in the 3 global TB watchlist countries, 2000&#8211;2020
<span class="subhead">Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=2.7, fig.width=8, fig.align='left', fig_2.2.7b, fig.alt="Trends in the estimated absolute number of TB deaths (HIV-positive and HIV-negative TB) in the 3 WHO watch-list countries, 2000−2020"} 

dta <- est[iso3 %in% c('KHM','RUS','ZWE')]
dta[, mort.milestone := mort.num[16] * 0.65, by = iso3]

  qplot(
    year,
    mort.num / 1e3,
    data = dta,
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_hline(aes(yintercept = mort.milestone / 1e3), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
    facet_wrap(~country, nrow=1, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
  xlab('') + ylab('TB deaths (total, in thousands)\nper year (log scale)') +
  expand_limits(y = 0) +
  scale_y_log10() +
  theme_gtb()


```




 



`r anch("Fig. 2.2.8")` 

### <span style="color:#F21905">Fig. 2.2.8</span> Estimated TB mortality rates in HIV-negative people, 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.2.8, fig.alt="Estimated TB mortality rates in HIV-negative people, 2020"} 

dta <- est[year == yr]
dta$var <- cut(
  dta$mort.nh,
  c(0, 1, 5, 20, 40, Inf),
  c('0\u20130.9', '1\u20134.9', '5\u201319', '20\u201339', '\u226540'),
  right = F,
  ordered_result = T
)

whomap(
  X = dta,
  legend.title = "Mortality\nper 100 000\npopulation per year",
  legend.pos = c(0.14, 0.34),
  colours = brewer.pal(5, "RdPu"),
)


```





 
 
 
 


`r anch("Fig. 2.2.9")`

### <span style="color:#F21905">Fig. 2.2.9</span> Estimates of the case fatality ratio (CFR), including HIV-negative and HIV-positive people, 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.2.9, fig.alt="Case fatality ratio (2020)"} 

dta <-
  est[year == yr & inc > 0, .(iso3,
                        cfr = 100 * mort / inc)]
dta$var <- cut(dta$cfr, 
               breaks=c(0, 5, 10, 15, 20, Inf),
               labels=c('0\u20134.9','5\u20139.9','10\u201314','15\u201319','\u226520'))

whomap(
  X = dta,
  colours = brewer.pal(5, "YlOrRd"),
  legend.title = 'Case fatality ratio\n(%)',
)

```






`r anch("Fig. 2.2.10a")`

### <span style="color:#F21905">Fig. 2.2.10a</span> Global distribution of estimated TB mortality in HIV-negative people by age group and sex (female in <span style="color:#951b81">purple</span>; male in <span style="color:#80a51b">green</span>), 2020 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=6, fig_2.2.10a, fig.alt="Main methods used to estimate TB mortality in HIV-negative people"} 

# source(here('disaggregation/reportoutput/plotfunctions.R'))
muplot <- function(indat,cnm){
    agz2 <- c('0_4','5_14','15_24','25_34','35_44',
              '45_54','55_64','65plus') #for labels
    agz3 <- gsub('_','-',agz2)
    agz4 <- c(rev(rev(agz3)[-1]),"\u226565")
    if(!'age' %in% names(indat)) indat[,age:=gsub("_","-",age_group)]
    indat[age=='65plus',age:=rev(agz4)[1]]
    indat$age <- factor(indat$age,levels=agz4,ordered=TRUE)
    indat$sex <- factor(indat$sex,levels=c('F','M'),ordered=TRUE)
    ## plot
    mp <- prodplot(data=indat,
                   mort ~ sex + age,
                   divider=mosaic()) +
        aes(fill=sex)
    mp <- mp + theme(axis.line=element_blank(),
                     axis.text.x=element_text(angle=90),
                     axis.text.y=element_text(),
                     axis.ticks=element_blank(),
                     axis.title.x=element_blank(),
                     axis.title.y=element_blank(),
                     legend.title=element_blank(),
                     legend.position="none",
                     panel.background=element_blank(),
                     panel.border=element_blank(),
                     panel.grid.major=element_blank(),
                     panel.grid.minor=element_blank(),
                     plot.background=element_blank()) +
        ggtitle(cnm)
    mp
}

disag.plot.mort.global <- function(D,title){
    ## ages and colors etc
    agz3 <- c('0\u20134','5\u201314','15\u201324','25\u201334','35\u201344','45\u201354',
              '55-64','65plus') #for labels
    agz4 <- c(rev(rev(agz3)[-1]),"\u226565")
    clz <- c('F'=palette_gtb('female'),
             'M'=palette_gtb('male'))
    ## plot construction
    plt <- muplot(D,title) +
        coord_flip() +
        scale_fill_manual(values=clz)+
        theme_gtb()+
        theme(axis.title.x=element_blank(), #remove
              axis.text.x=element_blank(),
              axis.ticks.x=element_blank(),
              legend.position = 'none')
    plt
}


annotate_figure(disag.plot.mort.global(Mglobsplt, 'Global'),
                left = text_grob('Age group (years)', rot = 90))



```








`r anch("Fig. 2.2.10b")`

### <span style="color:#F21905">Fig. 2.2.10b</span> Regional distribution of estimated TB mortality in HIV-negative people by age group and sex (female in <span style="color:#951b81">purple</span>; male in <span style="color:#80a51b">green</span>), 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=10, fig_2.2.10b, fig.alt="Main methods used to estimate TB mortality in HIV-negative people"} 

disag.plot.mort.regional <- function(D){
    ## ages and colors etc
    agz3 <- c('0\u20134','5\u201314','15\u201324','25\u201334','35\u201344','45\u201354',
              '55-64','65plus') #for labels
    agz4 <- c(rev(rev(agz3)[-1]),"\u226565")
    # whozt <- c('Africa','The Americas','Eastern Mediterranean','Europe',
    #            'South-East Asia',
    #            'Western Pacific')
    # regs <- c('AFR','AMR','EMR','EUR','SEA','WPR')
    clz <- c('F'=palette_gtb('female'),
             'M'=palette_gtb('male'))
    wr <- levels(D$name)
    ## plot construction
    Pltlst <- list()
    for(i in 1:6){
#        plt <- muplot(D[g.whoregion==regs[i]],whozt[i]) +
        plt <- muplot(D[name==wr[i]],wr[i]) +
            theme(legend.position="none")
#        if(!(i%%3==1)) plt <- plt + theme(axis.text.y=element_blank())
        plt <- plt + coord_flip() +
            scale_fill_manual(values=clz)+
            theme_gtb()+
            theme(axis.title.x=element_blank(),
                  axis.text.x=element_blank(),
                  axis.ticks.x=element_blank(),
                  legend.position = 'none')
        Pltlst[[i]] <- plt
    }
    p <- ggpubr::ggarrange(plotlist = Pltlst, ncol = 3,nrow=2)
    annotate_figure(p, left = text_grob('Age group (years)', rot=90))
}

disag.plot.mort.regional(Mregsplt)

```








`r anch("Fig. 2.2.11a")`
  
### <span style="color:#F21905">Fig. 2.2.11a</span> Main methods used to estimate TB mortality in HIV-negative people up to 2019

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.2.11a, fig.alt="Main methods used to estimate TB mortality in HIV-negative people up to 2019"} 

dta <-
  est[year == yr - 1, .(iso3,
                        source.mort)]
dta$var <- dta$source.mort

whomap(
  X = dta,
  legend.title = 'Main method',
  colours = c('#0a6acd', 'deepskyblue2', '#2BA81D')
)


```
<div class="footnote"> IHME: Institute of Health Metrics and Evaluation, Seattle, USA. VR: Vital registration of causes of deaths.</div> 







`r anch("Fig. 2.2.11b")`

### <span style="color:#F21905">Fig. 2.2.11b</span> Main methods used to estimate TB mortality in HIV-negative people in 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.2.11b, fig.alt="Main methods used to estimate TB mortality in HIV-negative people in 2020"} 

dta <-
  est[year == yr, .(iso3,
                    source.mort)]
dta$var <- factor(dta$source.mort, labels=c('Pre-2020 trend','Dynamic model','Statistical model','VR'))

whomap(
  X = dta,
  legend.title = 'Main method',
  colours = c('#fdae61', '#abd9e9', '#e5e5ff', '#2BA81D')
)

```

<div class="footnote"> VR: Vital registration of causes of deaths.</div> 







 
