--- 
title: "Chapter_2_1_text" 
author: "Philippe Glaziou, Pete Dodd" 
date: "`r Sys.Date()`" 
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
--- 





```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(echo = FALSE,  
                      results = "asis", 
                      message = FALSE, 
                      warning = FALSE) 
library(data.table)
library(gtbreport)
library(here)
library(kableExtra)
library(gridExtra)
library(whomap)
library(ggrepel)
library(ggpubr)
library(RColorBrewer)

load(here('report/data/gtb.rda'))

``` 






`r anch("Table 2.1.1")`


### <span style="color:#F21905">Table 2.1.1</span> Global and regional estimates of TB incidence, numbers (in thousands) and rates (per 100 000 population) in 2020
```{r tab_2.1.1}

tab.header <- c('Region or country group','Population','Best estimate','Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High'
                  )

knitr::kable(cbind(tabinc[c(32:37, 31, 38), 1:8], tab1b[c(32:37, 31, 38), 2:7]),
             format = "html",
             align = 'rccccccccccccc',
             col.names = tab.header,
             # Add a table ID so that it can be styled using extra CSS in Sitefinity
             table.attr = "id='burden_num_table'") |>
add_header_above(header = c(" " = 2, "Total" = 3, "HIV-positive" = 3, "Total" = 3, "HIV-positive" = 3)) |>
add_header_above(header = c(" " = 2, "Number of cases (in thousands)" = 6, "Rate per 100 000 population" = 6)) 
  # pack_rows("High burden countries", 1, 30) |>
  # pack_rows('Groupings', 31, 38) 


```
<div class="footnote">
Low and high are the 5th and 95th percentiles of the uncertainty interval (UI)
</div>






`r anch("Fig. 2.1.1")`

### <span style="color:#F21905">Fig. 2.1.1</span> Global trends in the estimated number of incident TB cases (left) and the incidence rate (right), 2000&#8211;2020
<span class="subhead">Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.1, fig.alt="Global trend in notifications of new and relapse TB cases, 2016-2020"} 


p1 <-
  qplot(
    year,
    inc.num / 1e6,
    data = global,
    geom = 'line',
    colour = I('blue')
  ) +
  geom_ribbon(
    aes(year, ymin = inc.lo.num / 1e6, ymax = inc.hi.num / 1e6),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, inc.h.num / 1e6), colour = I('lightskyblue2')) +
  geom_line(aes(year, c.newinc / 1e6)) +
  geom_ribbon(
    aes(year, ymin = inc.h.lo.num / 1e6, ymax = inc.h.hi.num / 1e6),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  ylab('Millions per year (log scale)') + xlab('') +
  annotate(
    'text',
    x = 2000,
    y = 9.5,
    label = 'All TB cases',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2005,
    y = 4.2,
    label = 'Notifications of new \nand relapse cases',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2000,
    y = 0.9,
    label = 'HIV-positive TB cases',
    hjust = 0,
    size = I(4)
  ) +
  expand_limits(y = 0) +
  scale_y_log10() +
  theme_bw(base_size = 14)

p2 <- qplot(year,
            inc,
            data = global,
            geom = 'line',
            colour = I('blue')) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, inc.h), colour = I('lightskyblue2')) +
  #  geom_line(aes(year, newinc)) +
  geom_line(aes(year, newinc)) +
  geom_ribbon(aes(year, ymin = inc.h.lo, ymax = inc.h.hi),
              fill = I('lightskyblue2'),
              alpha = 0.4) +
  ylab('Rate per 100 000 population per year (log scale)') + xlab('') +
  expand_limits(y = 0) +
  geom_hline(aes(yintercept = inc[16] * 0.8), linetype = I(2)) +
  annotate(
    'text',
    x = 2005,
    y = 120,
    label = '2020 milestone',
    hjust = 0,
    size = I(4)
  ) +
  scale_y_log10() +
  theme_bw(base_size = 14)

mp <- grid.arrange(p1, p2, ncol = 2)



```








`r anch("Fig. 2.1.2")`

### <span style="color:#F21905">Fig. 2.1.2</span> Trends in estimated TB incidence rates by WHO region, 2000&#8211;2020
<span class="subhead">Total TB incidence rates are shown in <span style="color:#3232ff">blue</span> and incidence rates of HIV-positive TB are shown in <span style="color:#87ceeb">light blue</span>. The black solid lines show notifications of new and relapse cases for comparison with estimates of the total incidence rate. Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8,  fig_2.1.2, fig.alt="Trends in estimated TB incidence rates by WHO region, 2016-2020"} 

qplot(year,
      inc,
      data = regional,
      geom = 'line',
      colour = I('blue')) +
  geom_line(aes(year, inc.h), colour = I('lightskyblue2')) +
  geom_ribbon(aes(year, ymin = inc.h.lo, ymax = inc.h.hi),
              fill = I('lightskyblue2'),
              alpha = 0.4) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  facet_wrap(~ region, scale = 'free_y') +
  xlab('') + ylab('Incidence rate per 100 000 population per year (log scale)') +
  expand_limits(y=0) +
  theme_bw(base_size = 12) +
  #  scale_y_continuous(trans='log10', breaks=c(1,2,5,10,20,30,50,100,200,300)) +
  scale_y_log10() +
  theme_gtb() +
  theme(legend.position = 'none') 


```








`r anch("Fig. 2.1.3a")`

### <span style="color:#F21905">Fig. 2.1.3a</span> Trends in estimated TB incidence rates in the 30 high TB burden countries compared with notifications of new and relapse cases, 2000&#8211;2020
<span class="subhead">TB incidence rates are shown in <span style="color:#3232ff">blue</span>. The black solid lines show notifications of new and relapse cases for comparison with estimates of the total incidence rate. Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=12, fig_2.1.3a, fig.alt="Trends in estimated TB incidence rates in the 30 high TB burden countries compared with notifications of new and relapse cases, 2000−2020"} 


qplot(year,
      inc,
      data = hest,
      geom = 'line',
      colour = I('blue')) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
  facet_wrap(~country, nrow=6, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
  expand_limits(y = 0) +
  xlab('') + ylab('Incidence rate per 100 000 population per year (log scale)') +
  scale_y_log10() +
  theme_gtb()

```




`r anch("Fig. 2.1.3b")`

### <span style="color:#F21905">Fig. 2.1.3b</span> Trends in estimated TB incidence rates in the 3 global TB watchlist countries compared with notifications of new and relapse cases, 2000&#8211;2020
<span class="subhead">TB incidence rates are shown in <span style="color:#3232ff">blue</span>. The black solid lines show notifications of new and relapse cases for comparison with estimates of the total incidence rate. Shaded areas represent uncertainty intervals. The horizontal dashed line shows the 2020 milestone of the End TB Strategy.</span> 
```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=2.7, fig.width=8, fig.align='left', fig_2.1.3b, fig.alt="Trends in estimated TB incidence rates in the 3 WHO watchlist countries compared with notifications of new and relapse cases, 2000−2020"} 

dta <- est[iso3 %in% c('KHM','RUS','ZWE')]
dta[, inc.milestone := inc[16] * 0.8, by = iso3]

qplot(year,
      inc,
      data = dta,
      geom = 'line',
      colour = I('blue')) +
  geom_ribbon(aes(year, ymin = inc.lo, ymax = inc.hi),
              fill = I('blue'),
              alpha = 0.4) +
  geom_line(aes(year, newinc)) +
  geom_hline(aes(yintercept = inc.milestone), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
  facet_wrap(~country, nrow=1, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
  expand_limits(y = 0) +
  xlab('') + ylab('Incidence rate per 100 000\npopulation per year (log scale)') +
  scale_y_log10() +
  theme_gtb()

```






`r anch("Fig. 2.1.4")`

### <span style="color:#F21905">Fig. 2.1.4</span> Estimated TB incidence in 2020, for countries with at least 100 000 incident cases
<span class="subhead">The eight countries that rank first to eighth in terms of numbers of cases, and that accounted for two thirds of global cases in 2020, are labelled.</span>


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.4, fig.alt="Estimated TB incidence in 2020, for countries with at least 100 000 incident cases"} 


dta <- est[year == yr & inc.num > 1e5, .(iso3, size = inc.num)]
dta <- merge(dta, cty[,.(iso3, country)], by='iso3')

bubblemap(dta, legend.pos = c(0.14, 0.5), legend.title = 'Number of\nincident cases',
          bubble.col = "#08519C",
          scale.limits = c(100000, 2600000),
          scale.breaks = c(100000, 500000,1000000,2000000),
          scale.labels = c('100 000', '500 000','1 000 000', '2 000 000')) +
  annotate(geom='text', label='India', x=80, y=-5) +
  annotate(geom='text', label='Pakistan', x=65, y=0) +
  annotate(geom='text', label='China', x=135, y=29, hjust=0) +
  annotate(geom='text', label='Bangladesh', x=135, y=23, hjust=0) +
  annotate(geom='text', label='Philippines', x=147, y=17, hjust=0) +
  annotate(geom='text', label='Indonesia', x=90, y=-15, hjust=0) +
  annotate(geom='text', label='Nigeria', x=-20, y=-5, hjust=0) +
  annotate(geom='text', label='South Africa', x=10, y=-45, hjust=0) +
  geom_segment(x=80, xend=80, y=-3, yend=20) +
  geom_segment(x=65, xend=65, y=3, yend=28) +
  geom_segment(x=110, xend=132, y=29, yend=29) +
  geom_segment(x=92, xend=132, y=23, yend=23) +
  geom_segment(x=122, xend=145, y=17, yend=17) +
  geom_segment(x=105, xend=115, y=-12, yend=-3) +
  geom_segment(x=-8, xend=8, y=-2, yend=9) +
  geom_segment(x=25, xend=25, y=-42, yend=-30) 

```





`r anch("Fig. 2.1.5")`

### <span style="color:#F21905">Fig. 2.1.5 </span>Estimated TB incidence rates, 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.5, fig.alt="Estimated TB incidence rates, 2020"} 


dta <- est[year==yr, .(iso3, inc)]
dta$var <- cut(
  dta$inc,
  c(0, 10, 50, 100, 300, 500, Inf),
  c('0-9.9', '10\u201349', '50\u201399', '100\u2013299', '300\u2013499', '\u2265500'),
  right = F,
  ordered_result = T
)

whomap(
  X = dta,
  legend.title =  "Incidence\nper 100 000\npopulation per year",
  legend.pos = c(0.14, 0.34)
)

```






`r anch("Fig. 2.1.6")`

### <span style="color:#F21905">Fig. 2.1.6</span> Estimated HIV prevalence in new and relapse TB cases, 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.6, fig.alt="Estimated TB incidence rates, 2020"} 


dta <- est[year==yr, .(iso3, tbhiv)]
dta$var <- cut(
  dta$tbhiv * 100,
  c(0, 5, 10, 20, 50, Inf),
  c('0\u20134.9', '5\u20139.9', '10\u201319', '20\u201349', '\u226550'),
  right = F,
  ordered_result = T
)

whomap(
  X = dta,
  colours = brewer.pal(5, "Reds"),
  legend.title =  "HIV prevalence,\nall ages (%)"
  ) 

```




`r anch("Fig. 2.1.7a")`
 
### <span style="color:#F21905">Fig. 2.1.7a</span> Global estimates of TB incidence (black outline) and case notifications disaggregated by age and sex (female in <span style="color:#951b81">purple</span>; male in <span style="color:#80a51b">green</span>), 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.7a, fig.alt="Global estimates of TB incidence (black outline) and case notifications disaggregated by age and sex (female in purple; male in green), 2020"} 

absspace <- function(x,...) {             #works
    format(abs(x), ..., big.mark=" ",scientific = FALSE, trim = TRUE)
}

disag.plot.inc.global <- function(D) {
  ## ages and colors etc -- From Pete, with a minor edit
  agz3 <- c('0\u20134',
            '5\u201314',
            '15\u201324',
            '25\u201334',
            '35\u201344',
            '45\u201354',
            '55\u201364',
            '65plus') #for labels
  agz4 <- c(rev(rev(agz3)[-1]), "\u226565")
  clz <- c('F' = palette_gtb('female'),
           'M' = palette_gtb('male'))
  ## plot construction
  plt <- ggplot(data = D, aes(x = age, y = newrel, fill = sex)) +
    coord_flip() +
    geom_bar(stat = 'identity', aes(y = ifelse(sex == 'M',
                                               newrel,
                                               -newrel))) +
    scale_fill_manual(values = clz) +
    geom_bar(stat = 'identity',
             fill = 'transparent',
             col = 1,
             aes(y = ifelse(sex == 'M',
                            inc,
                            -inc))) +
    theme_gtb() +
    theme(legend.position = "none") + xlab('Age group (years)') + ylab('') +
    scale_x_discrete(labels = agz4) +
    scale_y_continuous(labels = absspace)
  plt
}

disag.plot.inc.global(globsplt)


```




`r anch("Fig. 2.1.7b")`

### <span style="color:#F21905">Fig. 2.1.7b</span> Regional estimates of TB incidence (black outline, in thousand) and case notifications (in thousand) disaggregated by age and sex (female in <span style="color:#951b81">purple</span>; male in <span style="color:#80a51b">green</span>), 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=10, fig_2.1.7b, fig.alt="Regional estimates of TB incidence (black outline, in thousand) and case notifications (in thousand) disaggregated by age and sex (female in purple; male in green), 2020"} 


# disag.plot.inc.regional(regsplt)

agz3 <- c('0\u20134',
          '5\u201314',
          '15\u201324',
          '25\u201334',
          '35\u201344',
          '45\u201354',
          '55\u201364',
          '65plus') #for labels
agz4 <- c(rev(rev(agz3)[-1]), "\u226565")
clz <- c('F' = palette_gtb('female'),
         'M' = palette_gtb('male'))

ggplot(data = regsplt, aes(x = age, y = newrel/1000, fill = sex)) +
  coord_flip() +
  geom_bar(stat = 'identity', aes(y = ifelse(sex == 'M',
                                             newrel/1000,
                                             -newrel/1000))) +
  scale_fill_manual(values = clz) +
  geom_bar(stat = 'identity',
           fill = 'transparent',
           col = 1,
           aes(y = ifelse(sex == 'M',
                          inc/1000,
                          -inc/1000))) +
  xlab('Age group (years)') + ylab('') +
  scale_x_discrete(labels = agz4) +
  scale_y_continuous(labels = absspace) +
    facet_wrap(
    ~ name,
    scales = 'free',
    ncol= 3,
    labeller = label_wrap_gen(width = 25)
  ) +
  theme_gtb() +
  theme(legend.pos = 'none')


```





`r anch("Fig. 2.1.8a")`

### <span style="color:#F21905">Fig. 2.1.8a</span> Main methods used to estimate TB incidence up to 2019

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.8a, fig.alt="Main methods used to estimate TB incidence up to 2019"} 

dta <-
  est[year == yr - 1, .(iso3,
                        source.inc)]
dta$var <- dta$source.inc

whomap(
  X = dta,
  legend.title = 'Main method',
  colours = c('#e7bcbc', '#d73030', '#08519C', '#2BA81D')
)

```




`r anch("Fig. 2.1.8b")`

### <span style="color:#F21905">Fig. 2.1.8b</span> Main methods used to estimate TB incidence in 2020

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_2.1.8b, fig.alt="Main methods used to estimate TB incidence in 2020"} 

dta <-
  est[year == yr, .(iso3,
                    source.inc)]
dta$var <- factor(dta$source.inc, labels=c('Case notifications,\nStandard adjustment','Pre-2020 trend','Dynamic model','Statistical model'))


whomap(
  X = dta,
  legend.title = 'Main method',
  colours = brewer.pal(4, "RdYlBu"),
)

```

