---
title: "Section 3.3 Treatment and treatment coverage (figures)"
author: "Hazim Timimi"
date: "`r Sys.Date()`"
output: 
  html_fragment:
    # Don’t include a table of contents
    toc: no
    # Set standard figure width to 12 inches
    fig_width: 12
    # Don’t write figure captions
    fig_caption: FALSE

---


```{r setup, include=FALSE}
# Set chunk options.
# Results "asis" is useful to output markdown from a function
# Suppress messages, warnings and also the ## at the beginning of printed text

knitr::opts_chunk$set(echo = FALSE, 
                      results = "asis",
                      message = FALSE,
                      warning = FALSE)

# Kill any attempt at using factors, unless we explicitly want them!
options(stringsAsFactors=FALSE)

# Load output packages ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
library(ggplot2)
library(dplyr)
library(scales)
library(RColorBrewer)
library(whomap)
library(gtbreport)
library(kableExtra)
library(here)

# load section data
source(here('report/ch3-3-dataprep.r'))

# Quick functions for callouts in the text to the figures/tables:

# 1. lnk() Creates a link from text to a named anchor)
lnk <- function(display_name){
  return(paste0('<span class="red">[', 
                display_name, 
                '](#', 
                gsub("[^a-zA-Z0-9]", "-", tolower(display_name)),
                ')</span>'))
}

# 2. anch() creates the named anchor to go above the figure or table
anch <- function(display_name){
  return(paste0('<a name="', 
                gsub("[^a-zA-Z0-9]", "-", tolower(display_name)),
                '"></a>'))
}

# rmarkdown::render(here::here("report/ch3-3-figs.rmd"), output_file = "ch3-3-figs.html", output_dir = here::here("report/html/"))

```


```{css, echo=FALSE}

/* Styles to make it easier to see in the html_fragment; this CSS can be included in the CSS widget of Sitefinity */
  
.section.level3 {
  padding-bottom: 1em;
  border-bottom: 1px solid #BCBCBC;
  margin-bottom: 2em;
}


.subhead, .footnote {
  font-size: 80%;
  font-weight: normal;
  margin-top: 0;
  padding-top: 0;
}

.red, .red a {
  color: #F21905; /* red text to show figure number */
}

/* Deaths averted table */

/* Recreating simple striped bootstrap table */
#deaths_averted_table {
  border-spacing: 0;
  border-collapse: collapse;
  margin-top: 1em;
  margin-bottom: 1em;
  /* Next two lines to allow horizontal scrolling on narrow screens */
  display: block;
  overflow-x: auto;
}

#deaths_averted_table th {
  border-bottom: 2px solid #DDDDDD;
  padding: 8px;
}

#deaths_averted_table td {
  border-top: 1px solid #DDDDDD;
  padding: 8px;
}

/* light gray for odd rows */
#deaths_averted_table tr:nth-child(odd) td {
  background-color: #F5F5F5;	
}

/* Bold for the final row with thick line above */
#deaths_averted_table tr:last-child td {
  border-top: 2px solid #DDDDDD;
  font-weight:bold;	
}

/* light gray when hovering over a row */
#deaths_averted_table tr:hover td {
  background-color: #DDDDDD;
}

/* Centre-align all column headings except for the first */
#deaths_averted_table th:not(:first-child) {
  text-align: center !important;
}

/* prevent numbers from wrapping in any of the columns */
#deaths_averted_table td {
  white-space: nowrap;
}


```


_Draft! Prepared `r Sys.Date()` using country-reported data CSV files dated `r csv_datestamp` and estimates files dated  `r csv_estimate_datestamp`._

`r anch("Fig. 3.3.1")`


### <span class="red">Fig. 3.3.1</span> Estimated TB treatment coverage (new and relapse patients as a percentage of estimated TB incidence) in `r report_year-1`, 30 high TB burden countries, WHO regions and globally


```{r fig_3.3.1, fig.alt="Forest plot of TB treatment coverage in 30 countries, regionally and globally", fig.asp=1}

f3.3.1_plot <- f3.3.1_data %>% 

  ggplot(aes(x=entity,
             y=c_cdr)) +
  geom_point() +
  
  labs(x="",
       y="Treatment coverage (%)") +
  
  geom_pointrange(aes(ymin=c_cdr_lo,
                      ymax=c_cdr_hi),
                  size=1,
                  colour="Darkgreen") +
  
  expand_limits(y=0) +
  coord_flip() +
  theme_gtb() 
  

print(f3.3.1_plot)

```

`r anch("Fig. 3.3.2")`


### <span class="red">Fig. 3.3.2</span> The ten countries with the largest gaps between notifications of new and relapse (incident) TB cases and the best estimates of TB incidence, `r report_year-1`^a^


```{r fig_3.3.2, fig.alt="Bubble map of difference between notifications and estimated incidence for 10 countries"}

# To help find centres of bubbles use
# filter(whomap:::centroid, id %in% f3.3.2_data$iso3) 

f3.3.2_plot <- f3.3.2_data %>%
  
  bubblemap(legend.title = "Size of gap",
            legend.pos = c(0.14, 0.5),
            bubble.col = "purple",
            scale.breaks = c(7e4, 5e5, 1e6),
            scale.limits = c(7e4, 1e6),
            scale.labels = c("70 000","500 000","1 000 000")) +
  
  annotate(geom='text', label='China', x=150, y=38, hjust=0, size=3) +
  geom_segment(x=104, xend=145, y=37, yend=37) +

  annotate(geom='text', label='Viet Nam', x=150, y=25, hjust=0, size=3) +
  geom_segment(x=106, xend=145, y=17, yend=25) +

  annotate(geom='text', label='Philippines', x=150, y=17, hjust=0, size=3) +
  geom_segment(x=121, xend=145, y=17, yend=17) +

  annotate(geom='text', label='Pakistan', x=50, y=0, hjust=0, size=3) +
  geom_segment(x=70, xend=70, y=30, yend=5) +

  annotate(geom='text', label='India', x=65, y=-5, hjust=0, size=3) +
  geom_segment(x=80, xend=80, y=23, yend=-5) +

  annotate(geom='text', label='Bangladesh', x=60, y=-10, hjust=0, size=3) +
  geom_segment(x=90, xend=90, y=24, yend=-8) +

  annotate(geom='text', label='Indonesia', x=80, y=-15, hjust=0, size=3) +
  geom_segment(x=114, xend=105, y=0, yend=-12) +

  annotate(geom='text', label='Nigeria', x=-30, y=0, hjust=0, size=3) +
  geom_segment(x=8, xend=-5, y=10, yend=0) +

  annotate(geom='text', label='Democratic\nRepublic\nof the Congo', x=-30, y=-20, hjust=0, size=3) +
  geom_segment(x=24, xend=0, y=-2, yend=-20) +

  annotate(geom='text', label='South Africa', x=50, y=-30, hjust=0, size=3) +
  geom_segment(x=25, xend=45, y=-30, yend=-30)
  

print(f3.3.2_plot)

```
<div class="footnote">^a^ The ten countries ranked in order of the size of the gap between notified cases and the best estimates of TB incidence in `r report_year-1` are: `r gsub("(Philip)|(Democrat)", "the \\1\\2", knitr::combine_words(f3.3.2_data$country, oxford_comma=FALSE))`.</div>


`r anch("Fig. 3.3.3")`


### <span class="red">Fig. 3.3.3</span> Estimated global number of incident HIV positive TB cases<span style="color:`r gtbreport::palette_gtb("inch")`;"> (red)</span> compared with the global number of notified new and relapse TB cases known to be HIV-positive (black) and the global number of TB patients started on antiretroviral therapy<span style="color:#277abe;"> (blue)</span>, 2004-`r report_year-1`  
<span class="subhead">Shaded area represents uncertainty intervals</span>



```{r fig_3.3.3, fig.alt="Line and ribbon chart of estimated TB/HIV incidence, number notified and number on antiretroviral therapy"}

f3.3.3_plot <- f3.3.3_data %>% 

  ggplot(aes(x=year, y=hivpos, ymin=0)) +
  geom_line(size=1) +
  
  geom_ribbon(aes(x=year, ymin=e_inc_tbhiv_num_lo, ymax=e_inc_tbhiv_num_hi),
              fill=gtbreport::palette_gtb("inch"),
              alpha=0.4) +
  geom_line(aes(year, e_inc_tbhiv_num),
            size=1,
            colour=gtbreport::palette_gtb("inch")) +

  geom_line(aes(year, art),
            size=1,
            colour="#277abe") +

  scale_x_continuous(name="Year",
                     breaks = seq(2004, report_year-1, by = 3)) +

  scale_y_continuous(name = "New and relapse cases per year (millions)",
                     labels = function(i){i/1e6}) +
  xlab("Year") +
  
  theme_gtb() +
  
  # Get rid of annoying x-axis line and ticks
  theme(axis.line.x = ggplot2::element_blank(),
        axis.ticks.x = element_blank())

print(f3.3.3_plot)

```


`r anch("Fig. 3.3.4")`


### <span class="red">Fig. 3.3.4</span> Estimated coverage of antiretroviral therapy for HIV-positive TB cases (HIV-positive TB patients on antiretroviral therapy as a percentage of the estimated incidence of HIV-positive TB) in 2020, 30 high TB/HIV burden countries, WHO regions and globally


```{r fig_3.3.4, fig.alt="Forest plot of ART treatment coverage for TB/HIV patients in 30 countries, regionally and globally", fig.asp=1}

f3.3.4_plot <- f3.3.4_data %>% 

  ggplot(aes(x=entity,
             y=c_art)) +
  geom_point() +
  
  labs(x="",
       y="Treatment coverage (%)") +
  
  geom_pointrange(aes(ymin=c_art_lo,
                      ymax=c_art_hi),
                  size=1,
                  colour="#ed6476") +
  
  expand_limits(y=0) +
  coord_flip() +
  theme_gtb() 
  
print(f3.3.4_plot)

```


`r anch("Fig. 3.3.5")`

### <span class="red">Fig. 3.3.5</span> Treatment outcomes for new and relapse TB cases in `r report_year-2`, WHO regions and globally  


```{r fig_3.3.5, fig.alt="Horizontal bar chart showing TB treatment outcomes for WHO regions and globally"}

f3.3.5_plot <- f3.3.5_data %>% 
  
  ggplot(aes(x=entity, 
             y=value, 
             fill = factor(outcome,
                           levels = c("Treatment success",
                                      "Failure",
                                      "Died",
                                      "Lost to follow-up",
                                      "Not evaluated")))) +
 
  geom_col(position = position_stack(reverse = TRUE)) +
  coord_flip() +

  theme_gtb() +
  
  scale_fill_manual("", 
                    values = c("Treatment success" = palette_gtb("succ"),
                               "Failure" = palette_gtb("fail"),
                               "Died" = palette_gtb("died"),
                               "Lost to follow-up" = palette_gtb("ltfu"),
                               "Not evaluated" = palette_gtb("neval"))) +
  
  labs(x="", y="Percentage of cohort") +

  geom_text(data=subset(f3.3.5_data, 
                        outcome=="Treatment success"),
            aes(label = round(value, digits = 0)),
            position = position_stack(reverse = TRUE), 
            size=3,
            hjust=1.5,
            color="white")

  

print(f3.3.5_plot)

```


`r anch("Fig. 3.3.6")`

### <span class="red">Fig. 3.3.6</span> Treatment outcomes for new and relapse TB cases globally 2012-`r report_year-2` 


```{r fig_3.3.6, fig.alt="Panel of 2 horizontal bar charts showing TB treatment outcomes globally by year since 2012 for TB and TB/HIV", fig.asp=0.6}

f3.3.6_plot <- f3.3.6_data %>%
  
  ggplot(aes(year,
             value,
             fill = factor(outcome,
                           levels = c("Treatment success",
                                      "Failure",
                                      "Died",
                                      "Lost to follow-up",
                                      "Not evaluated")))) +
 
  geom_col(position = position_stack(reverse = TRUE)) +

  facet_wrap( ~ subgroup, nrow = 2) +
  
  coord_flip() +

  theme_gtb() +

  scale_fill_manual("", 
                    values = c("Treatment success" = palette_gtb("succ"),
                               "Failure" = palette_gtb("fail"),
                               "Died" = palette_gtb("died"),
                               "Lost to follow-up" = palette_gtb("ltfu"),
                               "Not evaluated" = palette_gtb("neval"))) +

  labs(x="Year started on treatment", y="Percentage of cohort") +

  scale_x_continuous(breaks = seq(2012, report_year-2)) +

  geom_text(data=subset(f3.3.6_data,
                        outcome=="Treatment success"),
            aes(label = round(value, digits = 0)),
			            position = position_stack(reverse = TRUE),
            size=3,
            hjust=1.5,
            color="white")



print(f3.3.6_plot)

```


`r anch("Fig. 3.3.7")`

### <span class="red">Fig. 3.3.7</span> Treatment outcomes for new and relapse HIV positive TB cases in `r report_year-1`, WHO regions and globally  


```{r fig_3.3.7, fig.alt="Horizontal bar chart showing TB treatment outcomes in HIV-positive cases for WHO regions and globally"}

f3.3.7_plot <- f3.3.7_data %>%

  # Note: plotting code same as 3.4.1 -- build a re-usable function?
  ggplot(aes(x=entity, 
             y=value, 
             fill = factor(outcome,
                           levels = c("Treatment success",
                                      "Failure",
                                      "Died",
                                      "Lost to follow-up",
                                      "Not evaluated")))) +
 
  geom_col(position = position_stack(reverse = TRUE)) +
  coord_flip() +

  theme_gtb() +
  
  scale_fill_manual("", 
                    values = c("Treatment success" = palette_gtb("succ"),
                               "Failure" = palette_gtb("fail"),
                               "Died" = palette_gtb("died"),
                               "Lost to follow-up" = palette_gtb("ltfu"),
                               "Not evaluated" = palette_gtb("neval"))) +
  
  labs(x="", y="Percentage of cohort") +

  geom_text(data=subset(f3.3.7_data, 
                        outcome=="Treatment success"),
            aes(label = round(value, digits = 0)),
            position = position_stack(reverse = TRUE), 
            size=3,
            hjust=1.5,
            color="white")


print(f3.3.7_plot)

```


`r anch("Fig. 3.3.8")`

### <span class="red">Fig. 3.3.8</span> Treatment success rate for new and relapse TB cases in children aged 0-14 years in `r report_year-2`, WHO regions and globally^a^ 


```{r fig_3.3.8, fig.alt="Horizontal bar chart showing TB treatment success rates in children for WHO regions and globally"}

f3.3.8_plot <- f3.3.8_data %>%
  
  ggplot(aes(x = entity, y = c_tsr_014)) +

  geom_bar(stat = "identity", fill = palette_gtb("succ")) +

  coord_flip() +

  theme_gtb() +

  scale_y_continuous(name = "Percentage of cohort",
                     limits = c(0,100)) +

  labs(x="") +

  geom_text(aes(label = round(c_tsr_014, digits = 0)),
            position = position_stack(reverse = TRUE), 
            size=3,
            hjust=1.5,
            color="white")


print(f3.3.8_plot)

```
<div class="footnote">^a^ Data reported by `r f3.3.8_txt$countries` countries on outcomes for `r int_spacer(f3.3.8_txt$kids_coh)` children aged 0-14 years, equivalent to `r ftb(f3.3.8_txt$kids_outcome_pct)`% of the `r int_spacer(f3.3.8_txt$kids_notified)` cases among children aged 0-14 years that were notified in `r report_year-2`.</div> 



`r anch("Fig. 3.3.9")`

### <span class="red">Fig. 3.3.9</span> Treatment outcomes for new and relapse TB cases (absolute numbers), 2000-`r report_year-2`, globally and for WHO regions  


```{r fig_3.3.9_glob, fig.alt="Bar chart showing treatment outcomes in absolute numbers by year since 2000 globally"}

f3.3.9_plot_glob <- f3.3.9_data %>%

  filter(entity=="Global") %>%
  ggplot(aes(year, 
             value,
             fill = factor(outcome,
                           levels = c("Treatment success",
                                      "Failure/Died/Lost to follow-up",
                                      "Not evaluated")))) +
 
  geom_col(position = position_stack(reverse = TRUE)) +

  facet_wrap( ~ entity) +

  theme_gtb() +
  
  scale_fill_manual("",
                    values = c("Treatment success" = palette_gtb("succ"),
                               "Failure/Died/Lost to follow-up" = palette_gtb("fail"),
                               "Not evaluated" = palette_gtb("neval"))) +
  
  labs(x="", y="Number of cases (millions)") +

  scale_x_continuous(breaks = seq(2000, report_year-2,by=6)) +
  
  # remove the legend
  theme(legend.position="none")

print(f3.3.9_plot_glob)

```


```{r fig_3.3.9_reg, fig.alt="Panel of bar charts showing treatment outcomes in absolute numbers by year since 2000 for WHO regions"}

f3.3.9_plot_reg <- f3.3.9_data %>%

  filter(entity!="Global") %>%
  ggplot(aes(year, 
             value,
             fill = factor(outcome,
                           levels = c("Treatment success",
                                      "Failure/Died/Lost to follow-up",
                                      "Not evaluated")))) +
 
  geom_col(position = position_stack(reverse = TRUE)) +

  facet_wrap( ~ entity,
              scales = "free_y") +

  theme_gtb() +
  
  scale_fill_manual("",
                    values = c("Treatment success" = palette_gtb("succ"),
                               "Failure/Died/Lost to follow-up" = palette_gtb("fail"),
                               "Not evaluated" = palette_gtb("neval"))) +
  
  labs(x="", y="Number of cases (millions)") +

  scale_x_continuous(breaks = seq(2000, report_year-2,by=6))


print(f3.3.9_plot_reg)

```
<div class="footnote">^a^ Cohorts before 2012 included new cases only.</div> 


`r anch("Table 3.3.1")`

### <span class="red">Table 3.3.1</span> Cumulative number of deaths averted by TB and TB/HIV interventions 2000-`r report_year - 1` (in millions), globally and by WHO region

```{r table_3.1}

# Format numbers for the table
t3.3.1_data <- t3.3.1_data %>%
  mutate(across(where(is.numeric), ftb))

# Create a table object using kable
table_header <- c('WHO region',
                  'Best estimate',
                  'Uncertainty interval',
                  'Best estimate',
                  'Uncertainty interval',
                  'Best estimate',
                  'Uncertainty interval')

knitr::kable(t3.3.1_data,
             format = "html",
             col.names = table_header,
             align = 'lccccc',
             # Add a table ID so that it can be styled using extra CSS in Sitefinity
             table.attr = "id='deaths_averted_table'") %>% 
  add_header_above(header = c(" " = 1,
                              "HIV-negative people" = 2,
                              "HIV-positive people" = 2,
                              "Total" = 2))


```

