--- 
title: "Chapter_Executive commentary" 
author: "Philippe Glaziou, Pete Dodd" 
date: "`r Sys.Date()`" 
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
--- 

## Figures

# FIG. E.1 Global trends in the estimated number of incident TB cases (left) and 
# the incidence rate (right), 2000–2020. 




# FIG. E.2 Global trends in the estimated number of TB deaths (left)
# and the mortality rate (right), 2000-2020.



# FIG. E.3 Top causes of death worldwide in 2019



# FIG. E.4 Global estimates of TB incidence (black outline)
# and case notifications disaggregated by age and
# sex (female in purple; male in green), 2020 PD



# FIG. E.6 Estimated TB incidence rates, 2020
# Use categories and names from Table EA6.1



# FIG. E.8 Trends in estimated TB incidence rates by WHO region, 2000–2020. 



# FIG. E.9 High TB burden countries estimated to have achieved the
# End TB Strategy 2020 milestone of a 20% reduction in the
# TB incidence rate between 2015 and 2020. PG Extract selected countries from Fig. 2.1.3



# FIG. E.10 Trends in the estimated absolute number of TB deaths
# (HIV-positive and HIV-negative) by WHO region, 2000-
# 2020



# FIG. E.11 High TB burden countries estimated to have achieved the
# End TB Strategy 2020 milestone of a 35% reduction in the
# absolute number of TB deaths between 2015 and 2020.



# FIG. E.17 Trends in annual and monthly/quarterly TB notifications, 
# selected countries NB this is to show the impact of the COVID-19 pandemic 
# on TB detection, and explains the impact shown in the preceding figure. PG



# FIG E.18 Estimated Impact of the COVID-19 pandemic on TB
# incidence, globally and for selected countries, up to 2025
# Selected countries will be those for which modelling has
# been done (approx. 15)



# FIG E.19 Estimated impact of the COVID-19 pandemic on TB
# mortality, globally and for selected countries, up to 2025
# Selected countries will be those for which modelling has
# been done (approx. 15)



# FIG E.20 The relationship between GDP per capita and the
# prevalence of undernutrition, and TB incidence per
# 100 000 population



## Tables

# TABLE E.1 Global TB targets set in the SDGs, the End TB Strategy 
# and the political declaration of the UN high-level meeting on TB, 
# for the period up to the SDG deadline of 2030 Reproduce 2020 version 
# of Table 2.1 SENT TO SUE



# TABLE E.2
# Cumulative number of deaths averted by TB and TB/HIV interventions
# 2000–2020 (in millions), globally and by WHO region PG



# TABLE E.3
# Global estimates of the number of TB cases attributable to selected
# risk factors. To be done after first round of TB incidence estimates is available















 

```{r setup, include=FALSE} 
# Set chunk options. 
# Results "asis" is useful to output markdown from a function 
# Suppress messages, warnings and also the ## at the beginning of printed text 

knitr::opts_chunk$set(echo = FALSE,  
                      results = "asis", 
                      message = FALSE, 
                      warning = FALSE) 

# Set output folder -- can only do this if not using the  
# RStudio knit button and instead doing the following from the  
# command line: 
# rmarkdown::render("create_chapter_example.Rmd", output_dir = "c:/scratch/") 

# load libraries, functions and data to output using another script 
source("load_libraries_functions_data_chapter_x.r") 

 
# Set a variable todecide whether or not to include accordian blocks  
# in the output 

include_accordian_blocks <- FALSE 

# Functions to produce boilerplate HTML for Sitefinity accordian block 
accordian_start <- function(){ 
  ifelse(include_accordian_blocks,  
         return('<div><div data-class="more-content"><div data-class="more-content-inner">'), 
         return('')) 
} 


accordian_end <- function(){ 
  ifelse(include_accordian_blocks,  
         return('</div></div><div data-class="more-button"><div data-class="more-button-label" data-close="" data-open=""><text>Show less</text> <text>Show more</text></div></div></div>'), 
         return('')) 
} 
 

``` 

# Chapter 5 

## TB diagnosis and treatment 

### Figure 5.1: Notifications of TB cases (new and relapse cases, all forms) (black) compared with estimated TB incident cases (<span class="green_text">green</span>), 2000–`r report_year - 1`, globally and for WHO regions. <span class="subhead">Shaded areas represent uncertainty intervals.</span> 
 

`r accordian_start()` 

 

```{r fig_5.1, fig.alt="Panel plot of TB notifications by WHO region and globally"} 

# Plot data is already prepared in dataframe called f5.1_data 
# I put summary statistics from the commentary below the figure in 
# variables f5.1_global_inc, f5.1_global_inc_lohi, f5.1_global_newinc 

f5.1_data %>% 
  ggplot(aes(x=year, y=c_newinc_millions, ymin=0)) + 
  geom_line(size=1) + 
  geom_ribbon(aes(x=year,  
                  ymin=e_inc_num_lo_millions,  
                  ymax=e_inc_num_hi_millions), 
              fill=palette_standard("incidence"), 
              alpha=0.4) + 
  geom_line(aes(year, e_inc_num_millions), 
            size=1, 
            colour=palette_standard("incidence")) + 
 

  facet_wrap( ~ entity, ncol = 4, scales="free_y") + 
  scale_x_continuous(name="Year", 
                     breaks = c(2000, 2009, report_year-1)) + 
  scale_y_continuous(name = "Millions per year") + 
  theme_gtbr_2021() + 
  # Add a gray line over the x-axis so that all graphs have a line at the bottom 
  annotate("segment", x=-Inf,  
           xend=Inf,  
           y=-Inf,  
           yend=-Inf,  
           colour = "#BCBCBC") 

``` 


There is still a large global gap between the estimated number of incident cases (`r f5.1_global_inc` million, range `r f5.1_global_inc_lohi` million, in `r report_year - 1`) and the number of people newly diagnosed (`r f5.1_global_newinc` million in `r report_year - 1`), due to underreporting of detected cases and underdiagnosis (if people with TB cannot access health care or are not diagnosed when they do). 


`r accordian_end()` 

### Table 5.1: Notifications of TB, HIV-positive TB, MDR/RR-TB and XDR-TB cases, globally and for WHO regions, `r report_year - 1` 
 

`r accordian_start()` 
 

```{r tab_5.1} 

# Create a table from data is already prepared in dataframe called t5.1_data 
# format the numrical data  
t5.1_data <- t5.1_data %>% 
  mutate_at(vars(-entity), 
            rounder) 

# Add % symbol to pct fields 
t5.1_data <- t5.1_data %>% 
  mutate(newrel_pulm_conf_pct = paste0(newrel_pulm_conf_pct, "%"), 
         newrel_ep_pct = paste0(newrel_ep_pct, "%")) 

# Create a table object using kable 
table_header <- c('WHO region', 
                  'Total notified', 
                  'New and relapse^a^', 
                  'Pulmonary new and relapse number', 
                  'Pulmonary new and relapse bactariologically confirmed (%)', 
                  'Extrapulmonary new and relapse (%)', 
                  'HIV-positive new and relapse', 
                  'MDR/RR-TB', 
                  'XDR-TB^b^') 


knitr::kable(t5.1_data, 
             format = "html", 
             col.names = table_header, 
             align = 'lrrrccrrr', 
             # Add a table ID so that it can be styled using extra CSS in Sitefinity 
             table.attr = "id='notifs_table'") 

``` 

<div class="footnote">^a^ *New and relapse* includes cases for which the treatment history is unknown. It excludes cases that have been re-registered as *treatment after failure*, as *treatment after lost to follow up* or as *other previously treated* (whose outcome after the most recent course of treatment is unknown or undocumented).    

^b^ XDR-TB is MDR-TB plus resistance to a fluoroquinolone and an injectable agent.</div> 

`r accordian_end()` 

